Small Data Center using Raspberry Pi 2 for Video Streaming

### 2015-06-05 paper/paper.tex
changes

% This .tex source is an example which *does* use
% the .bib file (from which the .bbl file % is produced).
% REMEMBER HOWEVER: After having produced the .bbl file,
% and prior to final submission, you *NEED* to 'insert'
% your .bbl file into your source .tex file so as to provide
% ONE 'self-contained' source file.
%
% refers to the cls file being used
\documentclass{sig-alternate-br}
\usepackage{float}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{enumitem}
\usepackage{tabularx}
\usepackage[numbers]{natbib}
\setlength{\bibsep}{0.0pt}
\newcommand\todo[1]{
	\textcolor{green}{#1}
}
\restylefloat{figure}
\restylefloat{table}
\begin{document}

%
% --- Author Metadata here --- DO NOT REMOVE OR CHANGE 
%\conferenceinfo{13$^{th}$ Twente Student Conference on IT}{June 23$^{st}$, 2010, Enschede, The Netherlands.}
\CopyrightYear{2015} % Allows default copyright year (200X) to be over-ridden - IF NEED BE.
%\crdata{0-12345-67-8/90/01}  % Allows default copyright data (0-89791-88-6/97/05) to be over-ridden - IF NEED BE.
% --- End of Author Metadata ---

@@

\maketitle
\begin{abstract}
Cloud computing is a trend IT that customers move computing and data away from desktop and portable PCs into data centers. These data centers require a lot of power and cooling. Nowadays approximately 30\% of the data coming from these data centers is video streaming. The Raspberry Pi 2 is a low cost device that can be used in a cloud for video streaming. In this paper a further investigation in the Raspberry Pi 2 cloud is done. This research is a design research in which a design is made for a video streaming cloud consisting of Raspberry Pi's. In this research it turned out that the Raspberry Pi 2 has its hardware limits. The results is that it is not so useful inside a data center, but it is good for education purposes.  In this research it was possible to run different video streaming algorithms. The Raspberry Pi 2 turned out to be a great device for doing research into video streaming.
\end{abstract}

@@

\section{Introduction}
Today most of us have some data in the cloud. But despite the attention from the community, research and development of Cloud Computing services are still in its early days~\citep{tso:2013}. \newline
Cloud computing is a trend in IT that moves customers computing and data away from desktop and portable PCs into large data centers \citep{dikaiakos:2009}. In the future most internet users are expected to access internet services over lightweight portable devices requiring a lot of data bandwidth \cite{dikaiakos:2009}. This bandwidth usage causes bottlenecks. To prevent these bottlenecks people build data centers around the globe. On these data servers a lot of improvements can be made \cite{abrahamsson:2013,beloglazov:2010}. 
These data servers for example require a lot of space and cooling. There are now new technologies such as for example the powerful ARM processor. Many companies want to explore the possibilities of the Raspberry Pi 2 , because of its ARM processor and the price \cite{Pcextreme}. Some data centers already offer some cloud computing using the Raspberry Pi 2. \newline
A Raspberry Pi 2 has a power usage between the 3-5 Watt and is a small computer. A normal server has a power usage between the 75 and 250 watt~\cite{Pcextreme,beloglazov2012energy}. The low power consumption and its computing power could mean that it is better to use a Raspberry Pi 2 for specific small tasks that do not demand an entire server. For this reason research is done on the performance of the Raspberry Pi 2 in a small data center. \newline
A Raspberry Pi cloud can be the small data center for the future \cite{tso:2013}. The Raspberry is sold for 40 euro, this price makes it cheaper to do research in compared to a normal server. Building a cloud like this can be a cost effective scale model\cite{tso:2013}. It's a ideal testbed for testing distributed software.

The Netherlands have one million subscribers for Netflix \cite{volkskrant}. Netflix is a popular video streaming service that makes HD movies watching possible. For this Netflix makes use of a content distribution network (CDN). On the internet this is known as an on demand service \cite{Adhikari:2012}. Netflix makes use of MPEG-DASH a protocol that makes streaming over HTTP possible \cite{martin:2013}. The problem is that Netflix is responsible for  29,7\% of the peak downstream traffic in US \cite{Adhikari:2012}. Because of this downstream the two main providers Comcast and  AT\&T were limiting the downstream of Netflix. This caused a lot of criticism and a new law for net neutrality has been made as a solution for this criticism \cite{net-neutrality}. \newline
In the Netherlands video streaming services such as Youtube and Netflix have an increasing amount of users. The data streaming problems coming from these increasing amount of users make it interesting to do more research in data centers with video streaming. Most of the research in computer science nowadays happen on expensive large servers \cite{tso:2013}. For this it can be very useful to see if it is possible to do research on a small computer like a Raspberry Pi 2. The problem is that we do not know if video streaming in a cloud consisting of Raspberry Pi's is possible.

In this paper there is an investigation in how useful the Raspberry Pi 2 is in a data center with video streaming. To come to an answer the performance of the Raspberry Pi 2 is evaluated. For this the main research question is: 
\begin{center}
How well does the Raspberry Pi 2 perform in small scale data centers with video streaming? 
\end{center}

In order to come to a answer for the main question several sub-questions are researched. These sub-questions will altogether provide a answer to the main research question.

@@

This research is part literature study and a part of it is building a small Raspberry Pi cloud to analyze balancing techniques on the Raspberry Pi 2. For this research cloud video streaming is one of the specific aspects that is investigated. The first two research questions describe the state of the art. The last three sub-questions go into the technical research that investigates the performance and possibility of Raspberry Pi's for cloud video streaming. The five sub-questions are being looked at in the next sections. Followed by a final section for conclusions and identified areas for future work.

\section{Research Methods}
This research investigates a cloud computer consisting of Raspberry Pi's. The research method for this would be the Design Science research method. this is a method to solve field problems. This research uses the Design Science method proposed by Hevner~\cite{hevner:2007}. This design method consists of three parts. 
\begin{figure}[H]
	\centering 
	\includegraphics[width=0.4\textwidth]{Design_science.png}
	\caption{Design Science}
	\label{fig:design} %always place your label after your caption!
\end{figure}
The \textbf{relevance cycle} is important, because you first have to identify the problems and opportunities. The identification of these problems we are done in the next section. \newline The \textbf{design cycle} is where you actually build the design. This is highlighted in section \ref{sec:setupex}. \newline The \textbf{rigor cycle} is a cycle were you can get two types of additional knowledge. It provides past knowledge to ensure its invention and it's there to guarantee that the new product is a research contribution. Information about the past can be found in section \ref{sec:back}. The other part is in section \ref{sec:setupex}.

\section{Background}\label{sec:back}
\subsection{Small scale computing}
Small scale cloud computing is cloud computing with smaller computation amounts than normal \cite{cox:2014}. Most users that operate in the cloud buy the services instead of buying the hardware and having to set up the servers themselves. This service is called Infrastructure as a Service. To this specific cloud service belongs the online storage of data. The Amazon Elastic Compute Cloud (Amazon EC2) is one of the most widely used infrastructure platforms that these clients use \cite{hofer:2011}. This elasticity of computing resources is very important \cite{Miettinen:2010:EEM:1863103.1863107}. A lot of companies have their own virtual machines running in large data centers \cite{beloglazov:2010}. The advantage is that these instances can be scaled dynamically to the customer it's need. Using these instances makes easier management of the computers and varying workloads possible.

Several cloud services have a huge demand on the network. This high demand can sometimes make it possible that data centers are unreachable.
That is a reason why many companies have their own data centers. These data centers are small and not energy efficient. Another reason why companies have their own data center is because they don't want their data to be in somebody else his hands.  Netflix a cloud service is responsible for approximately 30\% of the peak downstream traffic in US \cite{Adhikari:2012, computer-networking}. To improve this there are several data patterns possible, it can for example make a huge different in sending data in small bursts or in one large burst.

A CDN makes use of small data centers. This is because in a CDN you can get your data that is from America from a more nearby storage place. For example a website may be hosted in America. This website is connected to a CDN and in this way its possible to get your data from a more nearby server that has a copy of that website. For example if you live in Amsterdam you can perhaps get your data from London where the CDN is located instead of America. In this way the latency is decreased and there is a shorter connection distance and you have faster load times.

Power Usage Efficiency (PUE) is nowadays becoming a more important aspect in cloud computing. Keeping the power usage low is one of the main challenges of cloud computing. The data centers now are only 12 till 20\% of the time operating, while all the servers inside are still using energy. Keeping these data centers up requires a lot of electrical power. Therefore we perhaps need to more Green Cloud Computing solutions that reduce the electrical power consumption and reduce the environmental impact of data centers~\cite{beloglazov2012energy}. For example with data centers that are using microservers. More information about this is in the next section. \newline 
With the increase of lightweight portable devices the energy efficiency becomes more important. For example can the data center do the graphic calculations needed to create a good video, however then more data needs to be send of the network. This will save energy, but will increase the data over the network. This means that a trade-off has to be made between local processing and computation offloading \cite{Miettinen:2010:EEM:1863103.1863107}. Computation offloading makes the data centers do the calculations, therefore making cloud computing more important.

\subsection{Microserver}
Microservers come to good use in small data centers \cite{microserver}.  They are smaller in size, computer power and consume less electricity. This makes them perfect for high traffic usage tasks.  Its cheaper to a small device than a big server inside a data center, so smaller microservers make a data center more dynamic. The ability to add and remove more easily microservers can save a lot of energy. It's also better for optimization. A micro server often has a different processor compared to a normal server. ARM is a processor a micro server might use. This processor does not have so much overhead compared to a Intel one. The Intel processor is better at complex tasks, while the Arm processor is better at small tasks \cite{microserver}. The difference in these processes is because of their design. The ARM has the possibility to optimize it for a special task. In the data center of the future more of these dedicated processors will be used.

\subsection{Streaming}\label{sec:stream}
Video streaming happens with TCP or HTTP. TCP sends frames to the client and in these frames is the video stored. Of course if the available TCP send rate is large enough then the video will play without any delays. But if the TCP send rate is less then needed then the video play will alternate between periods of continuous playout and periods of freezing~\cite{computer-networking}. Its very important that video freezing or audio playing faster then the video does not happen. This is why multimedia applications have high performance requirements. To make the video streaming possible the client will connect with video streaming service its load balancer. The load balancer will then point were which videos are stored. To show how the processes operate with one a other a UML sequence diagram is used \cite{Bernardi:2002:USD:584369.584376}. Here below is in a UML sequence diagram the flow displayed between the client and the video service. 
\begin{figure}[H]
	\centering 
	\includegraphics[width=0.4\textwidth]{VideoStreaming.jpg}
	\caption{Video stream}
	\label{fig:stream} %always place your label after your caption!
\end{figure}

As visible in the diagram you can see that there is a loop for the streaming. In this loop the client will keep requesting for more parts of the videos. Depending on the algorithm being used the amount of data requested by the client can fluctuate. For streaming a video HTTP streaming is often used. By using HTTP the client can correctly receive the video, if it receives too much data then its buffer for video playing can get full. To prevent throwing data away the server will then stop sending. To prevent throwing data away the client will have a not to large buffer. \newline
There is a important streaming technique called adaptive streaming. This streaming algorithm is often used for video on demand (VOD). It improves the experience for the end-user by adapting the video quality dynamically to the viewers network condition \cite{ffmpeg}. This techniques looks to the clients CPU possibility and it's network possibilities. It delivers small sized chunks at multiple bitrates. This techniques is used for HTTP Live Streaming (HLS). This makes use of RTMP streaming and HTTP. This way of streaming is released in 2014. It codifies video in the H.264 format.\newline
There is another special kind of HTTP streaming called DASH (Dynamic Adaptive Streaming over HTTP)~\cite{computer-networking}. In DASH the video is encoded into several different versions. If the bandwidth is high, then the client selects chunks from a high-rate version and the way around~\cite{computer-networking}. The possibility to switch between bit rate versions is in video streaming very important. To make a smooth transition between these versions possible there are intermediate versions so that the client will see less suddenly the change in video quality. In HTTP streaming the server stores the video with a different URL.

\subsection{Video streaming and load balancing}
For video streaming several load balancing techniques can be used. There are several choices that have to be made by choosing a load balancer for a service. These choices can for example be performance, reliability and features. There can be a non-uniform demand for different movies. One solution to overcome this problem would be to replicate the most popular movies. One obvious problem is then that it is expensive in terms of storage space required. To solve this you can use dynamic replication to solve the load demand~\cite{dan1996load}. By using this you move portions of a movie to less used storage devices. The allocation of these movies only happen when there is a to high increase on one specific storage device. In this way load imbalances can be prevented \cite{dan1996load}. \newline
Load balancers are generally grouped in two layers. Layer 4 (FTP,IP, UDP, TCP )and layer 7 the application layer. Layer 7 load balancers distribute request based upon data in the application layer. Here below are some protocols and languages used for this:
\begin{enumerate}[topsep=0pt,itemsep=-1ex,partopsep=1ex,parsep=1ex]
	\item HTTP, is the hypertext transfer protocol. In section \ref{sec:stream} are mentioned some usages of this protocol. A lot of video streaming services use HTTP streaming \cite{Adhikari:2012}.  
	\item RTMP is a real time messaging protocol ~\cite{rtmp}. It is developed for streaming data between a player and a server. It is encapsulated in HTTP to traverse firewalls. It gives more video player options compared to normal HTTP thus giving the user a better experience. A disadvantage compared to normal HTTP is that it is sensitive for data spikes. These data spikes can result in a overload of the buffer which results in a empty buffer, this can lead to a stop in the video playing. 
	\item Synchronized Multimedia Integration Language (SMIL) describes multimedia presentations \cite{smil}. SMIL refers to media objects by URLs, allowing them to be shared between presentations and stored on different servers for load balancing. 
\end{enumerate}

@@

Nowadays a lot of people watch their movies online using video streaming and this amount of people is increasing. The Netherlands have for example one million subscribers for Netflix \cite{volkskrant}. One thing that these users want is the on demand video streaming. They do not want to store the data itself and they want to have a wide choice of different videos. Netflix is a video streaming service that makes HD movies watching possible. This is known as a on demand service. For this Netflix makes use of a content distribution network (CDN). They make use of amazon its AWS, simpleDB , S3 and Cassandra for file storage \cite{Adhikari:2012}. Netflix makes use of MPEG-DASH a protocol that makes streaming over HTTP possible. To make this possible Netflix does several things in the cloud.

@@

\subsubsection{Video stream CDN}
Video streamer Netflix has  it's own CDN. This CDN is fully operational in 2015 and now they still use some other CDN's like Akamai. Reasons for having an own CDN is that the load balancing can be better improved. The CDN that Netflix has created is called Open Connect. To make their CDN possible they work together with the ISP providers. They use high performance HTTP delivery to get the content to the users. By having their own CDN they can get the video's faster to their users, because they can optimize their video stream algorithms for their own CDN. Besides this its possible that in this way the data going over the network can be decreased.

@@

There have been several cloud projects with the Raspberry Pi. \newline
One of these has been the supercomputer build by Southampton~\cite{cox:2014}. Here they build it with 64 raspberry Pi's . They used a Message Passing interface to communicate between the Raspberry Pi's. The research was done in order to see what the performance of a low-power high performance cluster is. \newline
The project of the university of Glasgow~\cite{tso:2013}. This data center has 56 Raspberry Pi's. It is build for research and education purposes for a cloud data center. They have a Hadoop running on their servers. \newline
There was another project called the Beowulf cluster~\cite{beowolf-setup}. This project was created for a PhD assignment. This cluster is build for collaboratively processing sensor data.  A Beowulf cluster is simply a collection of identical, commodity computer hardware based systems~\cite{beowolf-setup}. The big advantage of the Raspberry Pi in this project was that it is cheap and it does not need a cluster administrator to watch over everything you do.

@@

\begin{figure}[H]
	\centering 
	\includegraphics[width=0.4\textwidth]{Pi2ModB1GB_-comp.jpeg}
	\caption{Raspberry Pi 2 model B}
	\label{fig:raspberry} %always place your label after your caption!
\end{figure}
\begin{table}[H]
	\centering \caption{Specifications}
	\begin{tabular}{|c|c|} \hline
		Ethernet & 100 Mbps \\ \hline
		USB & 4 x USB 2.0 \\ \hline
		Video out & HDMI 1.4 \\ \hline
		Audio & 2 x analog \\ \hline
		CPU & 900MHz quad-core ARM Cortex-A7 \\ \hline
		card slot & Micro SD  \\ \hline
	\end{tabular}
	\label{tab:Specificaties}
\end{table}
The Raspberry Pi 2 (RPi 2) only consumes 3 watts, because of this it does not need active cooling. The RPi 2 does have a limited ethernet cable of only 100 Mbps. This can be a huge drawback when large amounts of data need to be distributed over the Network. It is however very useful for a investigation in load balancing for video streaming. A small network cable offers still room for analyzing the video stream. The USB 2.0 has not a huge capabilities to share the data from a external harddisk to a user, but it will be enough for this small simulation. A normal Seagate harddisks demands 10 Watt when it is actively streaming videos and when it's standby it has a consumption of 0.1  Watt.

@@

A lot of small computers that can be turned of, instead of a few big ones that have to be on all the time can be a big step in saving energy consumed by big data centers. In this part we will look into the possibilities of doing this with the Raspberry Pi. \newline
It is possible to fit a lot of Raspberry Pi's inside a data center. PCextreme has made it possible to fit 500 Pi's inside a single data rack \cite{Pcextreme}. For this special designed boards are needed. With a special board you can for example reboot them from a distance. By having a microserver like a Raspberry Pi in a data center the hardware can be owned by only the customer, and because its in a data center there are the advantages of the high speed internet. By owning the hardware the customer will have more control over the data in the privacy and reliability aspect.\newline In this research we only have 8 Pi's and these fit easily inside the rack. The cooling of the data rack is more than sufficient for all these Raspberry's. The power consumption of the RPi 2 is low as you can see in table \ref{tab:cpu} below. The RPi 2 is also easy to replace inside a data center and it can easily be turned on and of, this makes it suitable inside a server rack. A problem that will exist with the Raspberry Pi 2 inside the data center is that every time a Pi's have troubles a help team inside the data center has to fix this for its customer. It is harder to get  a fully automated re-installation as for example a vServer more easily does.

>>>>

% This .tex source is an example which *does* use
% the .bib file (from which the .bbl file % is produced).
% REMEMBER HOWEVER: After having produced the .bbl file,
% and prior to final submission, you *NEED* to 'insert'
% your .bbl file into your source .tex file so as to provide
% ONE 'self-contained' source file.
%
% refers to the cls file being used
\documentclass{sig-alternate-br}
\usepackage{float}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{enumitem}
\usepackage{tabularx}

\usepackage[numbers]{natbib}
\setlength{\bibsep}{0.0pt}
\newcommand\todo[1]{
	\textcolor{green}{#1}
}
\restylefloat{figure}
\restylefloat{table}
\begin{document}
\setcounter{page}{1}
\pagenumbering{arabic}
%
% --- Author Metadata here --- DO NOT REMOVE OR CHANGE 
\conferenceinfo{23$^{th}$ Twente Student Conference on IT}{June 22$^{st}$, 2015, Enschede, The Netherlands.}
\CopyrightYear{2015} % Allows default copyright year (200X) to be over-ridden - IF NEED BE.
%\crdata{0-12345-67-8/90/01}  % Allows default copyright data (0-89791-88-6/97/05) to be over-ridden - IF NEED BE.
% --- End of Author Metadata ---

@@

\maketitle
\begin{abstract}
Cloud computing is a IT trend  in which customers move computing and data away from computer and laptop into data centers. These data centers require a lot of power and cooling. Nowadays, approximately 30\% of the data coming from these centers is video streaming. The Raspberry Pi 2 is a low cost device that can be used in a cloud for video streaming. In this paper a further investigation in the Raspberry Pi 2 cloud is done. This research is a design research in which a design is made for a video streaming cloud consisting of Raspberry Pi's. In this research it turned out that the Raspberry Pi 2 has its hardware limits. The results is that it is not so useful inside a data center, but it is good for education purposes.  In this research it was possible to run different video streaming algorithms. The Raspberry Pi 2 turned out to be a great device for doing research into video streaming.
\end{abstract}

@@

\section{Introduction}
Today, most of us have some data in the cloud. Despite the attention from the community, research and development of Cloud Computing services are still in its early days~\citep{tso:2013}. \newline
Cloud computing is a trend in IT that moves customers computing and data away from computer and laptop into large data centers \citep{dikaiakos:2009}. In the future, most internet users are expected to access internet services over lightweight portable devices requiring a lot of data bandwidth \cite{dikaiakos:2009}. This bandwidth usage causes bottlenecks. To prevent these bottlenecks, data centers are build around the globe. These data centers can be greatly improved \cite{abrahamsson:2013,beloglazov:2010}. 
For example, these data centers require a lot of space and cooling. There are now new technologies such as for example the powerful ARM processor. Many companies want to explore the possibilities of the Raspberry Pi 2 , because of its ARM processor and the low price \cite{Pcextreme}. At the time of writing the Raspberry Pi 2 costs around 40 euro. Some data centers already offer some cloud computing using the Raspberry Pi 2 \cite{Pcextreme}. \newline
A Raspberry Pi 2 is a small computer and has a power usage between the 2 and 4 Watt. In contrast, a normal server has a power usage between the 75 and 250 watt~\cite{Pcextreme,beloglazov2012energy}. The low power consumption of the Raspberry Pi 2 and its computing power could mean that it is better to use a Raspberry Pi 2 for specific small tasks that do not demand an entire server. For this reason research is done on the performance of the Raspberry Pi 2 in a small data center. \newline
A Raspberry Pi cloud can be the small data center for the future \cite{tso:2013}. The Raspberry Pi 2 is cheaper in price compared to a more expensive normal server. The Raspberry Pi 2 can be an ideal testbed for testing distributed software.

In The Netherlands, one million people use Netflix \cite{volkskrant}. Netflix is a popular video streaming service that provides HD movies. For this Netflix makes use of a Content Distribution Network (CDN). Netflix makes use of MPEG-DASH a protocol for streaming over HTTP \cite{martin:2013}. The problem is that Netflix is responsible for  29,7\% of the peak downstream traffic in the US \cite{Adhikari:2012}, because of the traffic the two main providers Comcast and  AT\&T were limiting the downstream of Netflix. This caused a lot of criticism, to solve this a new law for net neutrality has been made \cite{net-neutrality}. \newline
In the Netherlands, video streaming services such as Youtube and Netflix have an increasing number of users. Nowadays, the data streaming problems are increasing, this makes it interesting to do more research in data centers with video streaming. Most of the existing research in computer science is performed on expensive large servers \cite{tso:2013}. It can be very useful to see if it is possible to do research on a small computer like a Raspberry Pi 2.

This paper investigates how useful the Raspberry Pi 2 is in a data center with video streaming. To come to a conclusion the performance of the Raspberry Pi 2 is evaluated by doing several benchmark tests. For this the main research question is: 
\begin{center}
How well does the Raspberry Pi 2 perform in small scale data centers with video streaming? 
\end{center}

To answer for the main question, several sub-questions are researched, which are:

@@

The five sub-questions are looked at in the next sections. Followed by a final section for conclusions and identified areas for future work.
The first two research question are a literature study into video streaming and cloud computing. and a part of it is  The last three sub-questions go into the technical research that investigates the performance and possibility of Raspberry Pi's for cloud video streaming. This is done by building a small Raspberry Pi cloud.

\section{Research Methods}
In the research the  aim is to build a Raspberry Pi cloud to measure the performance. This is achieved by building a Raspberry Pi cloud and doing benchmark test on this cloud. This research uses the Design Science method proposed by Hevner~\cite{hevner:2007}. The method is used to make an architecture for a Raspberry Pi cloud. This design science method consists of three cycles:
\begin{figure}[H]
	\centering 
	\includegraphics[width=0.4\textwidth]{Design_science.png}
	\caption{Design Science cycles \cite{hevner:2007}}
	\label{fig:design} %always place your label after your caption!
\end{figure}
\textbf{relevance cycle} Represents the interaction with the domain. Here the problems and opportunities are identified. The identification of these problems are done in the next section. \newline 
\textbf{design cycle} The most important cycle where the design is build. This is highlighted in section \ref{sec:setupex}. \newline
\textbf{rigor cycle} Represents the interaction with the scientific community. This information can be found in section \ref{sec:back} and \ref{sec:setupex}.

\section{Background}\label{sec:back}
\subsection{Small scale computing}
Small scale cloud computing is cloud computing with smaller computation amounts than normal \cite{cox:2014}. Most customers for cloud solutions buy the services instead of buying the hardware and having to set up the servers themselves. This service is called Infrastructure as a Service (IaaS). The elasticity of computing resources in the cloud is important to provide IaaS \cite{Miettinen:2010:EEM:1863103.1863107}. Many companies have their software running on virtual machines in data centers \cite{beloglazov:2010}. The advantage is that these virtual machines are scaled dynamically to the customer its need. Using the cloud makes  management easier.

Several cloud services have a huge network demand, this may cause data centers to be unreachable. Because of the privacy and reachability of a data center many companies have their own servers. These servers are not energy efficient. \newline
Netflix a cloud service is responsible for approximately 30\% of the peak downstream traffic in US \cite{Adhikari:2012, computer-networking}. The data flowing through the network has different patterns. To improve this there are several data patterns possible, it can for example make a huge different in sending data in small bursts or in one large burst \cite{computer-networking}.

Power Usage Efficiency (PUE) is a important aspect in cloud computing \cite{beloglazov2012energy}. There are more lightweight portable devices coming \cite{Miettinen:2010:EEM:1863103.1863107}, these devices are energy efficient, but will demand more processes being done by data centers \cite{Miettinen:2010:EEM:1863103.1863107}. Nowadays, the data centers operate 12 to 20\% of the time, the other 80 \% of the time the servers inside are still using energy \cite{beloglazov2012energy}. Therefore there is a need Green Cloud Computing solutions that reduce the electrical power consumption and the environmental impact of data centers~\cite{beloglazov2012energy}. Microservers can be used to reduce the consumption.

\subsection{Microserver}
Microservers are used in data centers \cite{microserver}. Microservers consume less electricity and have enough computer power for small tasks, therefore they can be used for high traffic usage tasks. Besides this they are small and are easier replaced compared to a normal server, by using them data centers can be more dynamic \cite{microserver}. The ability to add and remove microservers more easily can save a lot of energy. \newline
A microserver often has a different processor compared to a normal server. The processors are often microprocessors like a ARM processor \cite{microserver}. There are two big processor companies in the world called ARM, Intel and AMD. The Intel and AMD processors are better at complex tasks, while the ARM processor is better at small tasks \cite{microserver}. In the data center of the future more dedicated microservers will be used, they are better at specialized tasks and are more energy efficient \cite{microserver}.

\subsection{Video streaming}\label{sec:stream}
Video streaming happens with HTTP, HTTP sends frames to the client and in the frames is the video stored \cite{computer-networking}. If the send rate is large enough the video will play without any delays, but if the send rate is too less then the video play will alternate between periods of continuous playout and periods of freezing~\cite{computer-networking}. \newline 
To make video streaming possible the client will connect with video streaming service its load balancer. The load balancer will then locate the videos. To show how the processes and operators operate with each other a UML sequence diagram is used \cite{Bernardi:2002:USD:584369.584376}. 
\begin{figure}[H]
	\centering 
	\includegraphics[width=0.4\textwidth]{VideoStreaming.jpg}
	\caption{Video stream}
	\label{fig:stream} %always place your label after your caption!
\end{figure}

In the diagram there is a loop for the video streaming, here the client will keep requesting for more parts of the videos. Depending on the algorithm being used the amount of data requested by the client can fluctuate. For streaming a video HTTP streaming is often used. By using HTTP the client can correctly receive the video, if it receives too much data its buffer for video playing gets full. To prevent throwing data away the server then stops sending.

There is a important streaming technique called adaptive streaming. This streaming algorithm is often used for video on demand (VOD). It improves the experience for the end-user by adapting the video quality dynamically to the viewers network condition \cite{ffmpeg}. This techniques looks to the clients CPU possibility and its network possibilities. It delivers small sized chunks at multiple bitrates. This techniques is used for HTTP Live Streaming (HLS). This makes use of RTMP streaming and HTTP. This way of streaming is released in 2014. It codifies video in the H.264 format.\newline
There is another special kind of HTTP streaming called DASH (Dynamic Adaptive Streaming over HTTP)~\cite{computer-networking}. In DASH the video is encoded into several different versions. If the bandwidth is high, then the client selects chunks from a high-rate version and the way around~\cite{computer-networking}. The possibility to switch between bit rate versions is in video streaming very important. To make a smooth transition between these versions possible there are intermediate versions so that the client will see less suddenly the change in video quality. In HTTP streaming the server stores the video with a different URL.

\subsection{Video streaming and load balancing}
For video streaming several load balancing techniques are used. To choose a technique there is looked at performance, reliability and features. There can be a non-uniform demand for different movies. One solution to overcome this problem would be to replicate the most popular movies, but this is expensive in terms of storage space required for this. A dynamic replication can be used to solve the load demand~\cite{dan1996load}. By using this  portions of a movie are moved to less used storage devices, so that these can stream the movie. The allocation of these movies only happen when there is a to high increase on one specific storage device. In this way load imbalances can be prevented \cite{dan1996load}. \newline
Load balancers are generally grouped in two layers. Layer 4 (FTP,IP, UDP, TCP )and layer 7 the application layer. Layer 7 load balancers distribute request based upon data in the application layer. Here below are some protocols and languages used for this:
\begin{enumerate}[topsep=0pt,itemsep=-1ex,partopsep=1ex,parsep=1ex]
	\item HTTP, is the hypertext transfer protocol. In section \ref{sec:stream} are mentioned some usages of this protocol. A lot of video streaming services use HTTP streaming \cite{Adhikari:2012}.  
	\item RTMP is a real time messaging protocol ~\cite{rtmp}. It is developed for streaming data between a player and a server. It is encapsulated in HTTP to traverse firewalls. It gives more video player options compared to normal HTTP thus giving the user a better experience. A disadvantage compared to normal HTTP is that it is sensitive for data spikes. These data spikes can result in a overload of the buffer which results in a empty buffer, this can lead to a stop in the video playing. 
	\item Synchronized Multimedia Integration Language (SMIL) describes multimedia presentations \cite{smil}. SMIL refers to media objects by URLs, allowing them to be shared between presentations and stored on different servers for load balancing. 
\end{enumerate}

@@

One thing that users want is on demand video streaming. They do not want to store the data itself, and they want to have a wide choice of different videos when they want to watch it. Netflix is a on demand video streaming service that makes HD movies watching possible. For this Netflix makes use of a content distribution network (CDN). They make use of amazon its AWS, simpleDB , S3 and Cassandra for file storage \cite{Adhikari:2012}. Netflix makes use of MPEG-DASH a protocol that makes streaming over HTTP possible. To make this possible Netflix does several things in the cloud.

@@

\subsubsection{Video stream CDN}
A CDN makes use of small data centers. This is because users want their data as nearby as possible  In this way the latency is decreased and there is a shorter connection distance, this gives the user faster load times. Video streamer Netflix has  its own CDN. This CDN is fully operational in 2015 and now they still use some other CDN's like Akamai. Reasons for having an own CDN is that the load balancing can be better improved. The CDN that Netflix has created is called Open Connect. To make their CDN possible they work together with the ISP providers. They use high performance HTTP delivery to get the content to the users. By having their own CDN they can get the video's faster to their users, because they can optimize their video stream algorithms for their own CDN. Besides this its possible that in this way the data going over the network can be decreased.

@@

There have been several cloud projects with the Raspberry Pi. \newline
One of these has been the supercomputer build by Southampton~\cite{cox:2014}. Here they build it with 64 raspberry Pi's . They used a Message Passing interface to communicate between the Raspberry Pi's. The research was done in order to see what the performance of a low-power high performance cluster is. \newline
The project of the university of Glasgow~\cite{tso:2013}. This data center has 56 Raspberry Pi's. It is build for research and education purposes for a cloud data center. They have a Hadoop running on their servers. \newline
There was another project called the Beowulf cluster~\cite{beowolf-setup}. This project was created for a PhD assignment. This cluster is build for collaboratively processing sensor data.  A Beowulf cluster is simply a collection of identical, commodity computer hardware based systems~\cite{beowolf-setup}. The big advantage of the Raspberry Pi in this project was that it was cheap and it does not need a cluster administrator to check everything that was being done.

@@

\begin{figure}[H]
	\centering 
	\includegraphics[width=0.4\textwidth]{Pi2ModB1GB_-comp.jpeg}
	\caption{Raspberry Pi 2 model B}
	\label{fig:raspberry} %always place your label after your caption!
\end{figure}
\begin{table}[H]
	\centering \caption{Specifications}
	\begin{tabular}{|c|c|} \hline
		Ethernet & 100 Mbps \\ \hline
		USB & 4 x USB 2.0 \\ \hline
		Video out & HDMI 1.4 \\ \hline
		Audio & 2 x analog \\ \hline
		CPU & 900MHz quad-core ARM Cortex-A7 \\ \hline
		card slot & Micro SD  \\ \hline
	\end{tabular}
	\label{tab:Specificaties}
\end{table}
The Raspberry Pi 2 (RPi 2) only consumes 3 watts, because of this it does not need active cooling. The RPi 2 does have a limited ethernet cable of only 100 Mbps. This can be a huge drawback when large amounts of data need to be distributed over the Network. It is however very useful for a investigation in load balancing for video streaming. A small network cable offers still room for analyzing the video stream. The USB 2.0 has not a huge capabilities to share the data from a external harddisk to a user. A external Seagate harddisk demands 10 Watt when it is actively streaming videos and when it is standby it has a consumption of 0.1  Watt.

@@

A lot of small computers that can be turned of, instead of a few big ones that have to be on all the time can be a big step in saving energy consumed by big data centers. In this part we will look into the possibilities of doing this with the Raspberry Pi 2 . \newline
It is possible to fit a lot of Raspberry Pi's inside a data center. PCextreme has made it possible to fit 500 Pi's inside a single data rack \cite{Pcextreme}. For this special designed boards are needed. With a special board RPi 2 can be rebooted from a distance. By having a microserver like a Raspberry Pi in a data center the hardware can be owned by only the customer, and because its in a data center there are the advantages of the high speed internet. By owning the hardware the customer will have more control over the data in the privacy and reliability aspect.\newline In this research there are 3 Pi's and these fit easily inside the rack. The cooling of the data rack is more than sufficient for all these Raspberry's. The power consumption of the RPi 2 is low as seen in table \ref{tab:cpu}. The RPi 2 is easy to replace inside a data center and it can easily be turned on and off, this makes it suitable inside a server rack. A problem that will exist with the Raspberry Pi 2 inside the data center is that every time a Pi's have troubles a help team inside the data center has to fix this for its customer. It is harder to get  a fully automated re-installation as for example a vServer more easily does.

In this project a own cluster of Raspberry Pi's has been build to test its possibilities. The setup for this project looked like this:
\begin{figure}[H]
	\centering 
	\includegraphics[width=0.45\textwidth]{setup.jpg}
	\caption{Project Setup}
	\label{fig:projectsetup} %always place your label after your caption!
\end{figure}
The figure \ref{fig:projectsetup} is photo of the project setup. This project setup is comparable with figure \ref{fig:setup}. The numbers in the front pint to the Raspberry Pi's that are also showed in figure \ref{fig:setup}. The number in the back shows the little router that is behind the power supply. In this picture there is a extra switch to connect all the Raspberry Pi's. To redo this experiment there are only three Raspberry Pi's necessary. \newline 
This setup can be placed inside of a data center. The switch and the router will then be removed and the rest will be fit into the server rack. The setup that has been made now could easily fit on the project table, so that it was easy to reach them and a lot of research on them could be done.
